Exploring dimensions of biodiversity in Japanese ferns

### 2021-08-24 ms/manuscript.Rmd
Update MS
- Use CVD-friendly colors for maps
- Make coastline thinner to improve visibility
- Move color palette setup to ms_functions.R

```{r setup, include = FALSE}
# Load libraries used for rendering Rmds
library(flextable)
library(kableExtra)
library(ftExtra)
library(scales)
library(patchwork)

@@

# Load functions only used when rendering Rmds
source(here::here("R/ms_functions.R"), local = TRUE)
```

```{r load-data}
# Load objects from targets workflow
tar_load(c(
  "repro_data",
  "occ_point_data_ferns",
  "occ_point_data_ferns_unfiltered",
  "biodiv_ferns_spatial",
  "k_taxonomy",
  "k_phylogeny",
  "signif_cells_protected_area",
  "protected_areas",
  "japan_shp",
  "phy_sig_results",
  "binary_sig_results",
  "traits_summary",
  "fern_traits",
  "japan_pterido_tree",
  "japan_fern_tree",
  "ppgi",
  "occ_point_data_ferns",
  "comm_ferns_full",
  "comm_ferns",
  "japan_points_raw",
  "plastome_calibration_dates",
  "plastome_tree",
  "green_list",
  "model_params",
  "lrt_results",
  "aic_env_repro",
  "model_fits",
  "biodiv_ferns_cent",
  "refs_yaml",
  "refs_other_yaml",
  "template_file",
  "csl_file",
  "broad_alignment_list",
  "japan_rbcL"
))
```

@@

Sequences for the plastid, coding *rbcL* gene are available for ca. 97% of the Japanese fern flora [@Ebihara2019b]. 
However, using this dataset alone for phylogeny suffers two drawbacks: it is not densely sampled enough to resolve many internal nodes, and it lacks many lineages that have fossils available for molecular dating. 
Because one goal of the study is to understand the distribution of paleo- vs. neo-endemic areas (defined in units of time), we require an ultrametric tree. 
To obtain a maximally sampled, robust phylogenetic tree, we combined the *rbcL* data of @Ebihara2019b (`r n_japan_fern_taxa` taxa) with a globally sampled plastid dataset including all fern sequences on GenBank for four widely-sequenced plastid genes (*atpA*, *atpB*, *rbcL*, and *rps4*; `r n_ftol_fern_taxa` taxa) and `r n_plastome_genes` other coding, single-copy plastid genes extracted from `r n_plastome_fern_taxa` complete fern plastomes (Nitta et al., in prep.) (`r ape::Ntip(plastome_tree) %>% scales::number(big.mark = ",")` OTUs total, including outgroup taxa). 
We conducted phylogenetic analysis with `r software("IQTREE")` [@Nguyen2015] and molecular dating with `r software("treePL")` [@Smith2012] using `r nrow(plastome_calibration_dates)` fossil calibration points after @Testo2016a, with the exception of treating fossil *Kuylisporites* as belonging to crown *Cyathea*, not *Cyathea* + *Alsophila* [@Loiseau2020]. 
We then trimmed the tree to Japanese taxa only and used this for all further analyses.
Tree visualization was done with the `r pack("ggtree")` R package [@Yu2017].

@@

```{r plot-prep}
# Define nicely formatted terms for printing variable names
resp_var_lookup <- tribble(
  ~resp_var, ~resp_var_pretty,
  "fd_obs_z", "SES of FD",
  "pd_obs_z", "SES of PD",
  "pe_obs_p_upper", "PE *p*-value",
  "rfd_obs_z", "SES of RFD",
  "richness", "Richness",
  "rpd_obs_z", "SES of RPD",
)

@@

# Crop high-res map of Japan to spatial div results area
japan <- sf::st_crop(japan_shp, sf::st_bbox(biodiv_ferns_spatial))

# Set CRS
japan_crs <- sf::st_crs(japan)
biodiv_ferns_spatial <- sf::st_set_crs(biodiv_ferns_spatial, japan_crs)

# Set color palettes:
# - Randomization test significance
signif_cols <-
  c(
    "> 0.99" = "#e31a1c", 
    "> 0.975" = "#fb9a99",
    "< 0.01" = "#1f78b4", 
    "< 0.025" = "#a6cee3",
    "not significant" = "#ffffcc"
  )
# - CANAPE
canape_cols <-
  c(
    "neo" = "#e31a1c", 
    "paleo" = "#1f78b4",
    "mixed" = "#6a51a3",
    "super" = "#4a1486",
    "not significant" = "#ffffcc"
  )
# - Bioregions
bioregion_cols <- rcartocolor::carto_pal(9, "Bold")[1:8] %>%
  set_names(1:8)
```

@@

```{r raw-diversity-make}
# Richness
a <- ggplot() +
  geom_sf(data = japan, fill = "grey90", color = "black", lwd = 0.25) +
  geom_sf(data = biodiv_ferns_spatial, aes(fill = richness), color = "transparent", alpha = 0.8) +
  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**(a)** Richness</span>", hjust = "inward", vjust = "inward", color = "white") + 
  labs(fill = "No. taxa") +
  scale_fill_scico(palette = "lajolla") + 
  scale_x_continuous(breaks = c(125, 135, 145)) +
  scale_y_continuous(breaks = c(25, 35, 45)) +
  map_theme()

# Raw PD
b <- ggplot() +
  geom_sf(data = japan, fill = "grey90", color = "black", lwd = 0.25) +
  geom_sf(data = biodiv_ferns_spatial, aes(fill = pd_obs), color = "transparent", alpha = 0.8) +
  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**(b)** PD</span>", hjust = "inward", vjust = "inward", color = "white") +
  labs(fill = "Relative\nbranch\nlength") +
  scale_fill_scico(palette = "lajolla") + 
  scale_x_continuous(breaks = c(125, 135, 145)) +
  scale_y_continuous(breaks = c(25, 35, 45)) +
  map_theme() +
  theme(
    axis.text.y = element_blank()
  )

# Raw FD
c <- ggplot() +
  geom_sf(data = japan, fill = "grey90", color = "black", lwd = 0.25) +
  geom_sf(data = biodiv_ferns_spatial, aes(fill = fd_obs), color = "transparent", alpha = 0.8) +
  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**(c)** FD</span>", hjust = "inward", vjust = "inward", color = "white") + 
  labs(fill = "Relative\nbranch\nlength") +
  scale_fill_scico(palette = "lajolla") + 
  scale_x_continuous(breaks = c(125, 135, 145)) +
  scale_y_continuous(breaks = c(25, 35, 45)) +
  map_theme() +
  theme(
    axis.text.y = element_blank()
  )

raw_diversity_plot <- a + b + c +
  plot_layout(nrow = 1)
```

@@

```{r rand-diversity-make}
# PD
a <- ggplot() +
  geom_sf(data = japan, fill = "grey90", color = "black", lwd = 0.25) +
  geom_sf(data = biodiv_ferns_spatial, aes(fill = pd_signif), color = "transparent", alpha = 0.8) +
  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**(a)** PD</span>", hjust = "inward", vjust = "inward", color = "white") + 
  scale_fill_manual(values = signif_cols) +
  map_theme() +
  theme(
    axis.text.x = element_blank(),
    legend.title = element_blank()
  )

# RPD
b <- ggplot() +
  geom_sf(data = japan, fill = "grey90", color = "black", lwd = 0.25) +
  geom_sf(data = biodiv_ferns_spatial, aes(fill = rpd_signif), color = "transparent", alpha = 0.8) +
  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**(b)** RPD</span>", hjust = "inward", vjust = "inward", color = "white") + 
  labs(fill = "") +
  scale_fill_manual(values = signif_cols) +
  map_theme() +
  theme(
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    legend.title = element_blank()
  )

# FD
c <- ggplot() +
  geom_sf(data = japan, fill = "grey90", color = "black", lwd = 0.25) +
  geom_sf(data = biodiv_ferns_spatial, aes(fill = fd_signif), color = "transparent", alpha = 0.8) +
  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**(c)** FD</span>", hjust = "inward", vjust = "inward", color = "white") + 
  labs(fill = "") +
  scale_fill_manual(values = signif_cols) +
  scale_x_continuous(guide = guide_axis(check.overlap = TRUE)) +
  map_theme() +
  theme(legend.title = element_blank())

# RFD
d <- ggplot() +
  geom_sf(data = japan, fill = "grey90", color = "black", lwd = 0.25) +
  geom_sf(data = biodiv_ferns_spatial, aes(fill = rfd_signif), color = "transparent", alpha = 0.8) +
  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**(d)** RFD</span>", hjust = "inward", vjust = "inward", color = "white") + 
  labs(fill = "") +
  scale_fill_manual(values = signif_cols) +
  scale_x_continuous(guide = guide_axis(check.overlap = TRUE)) +
  map_theme() +
  theme(
    axis.text.y = element_blank(),
    legend.title = element_blank()
  )

@@

# Generate final plot
modelfit_plot_list <- 
  model_fits %>%
  mutate(
    signif_var = str_replace_all(resp_var, "_obs_z", "_signif"),
    resp_var_print = case_when(
      resp_var == "richness" ~ "Richness (no. spp.)",
      TRUE ~ str_to_upper(resp_var) %>% word(1, 1, "_") %>% paste("SES of", .)
    ),
    model_data = list(biodiv_ferns_used_for_model)) %>%
  left_join(model_subplot_order, by = c("resp_var", "indep_var")) %>%
  arrange(order) %>%
  pmap(plot_fits)

@@

```{r endemism-make}
endemism_plot <- ggplot() +
  geom_sf(data = japan, fill = "grey90", color = "black", lwd = 0.25) +
  geom_sf(data = biodiv_ferns_spatial, aes(fill = endem_type), color = "transparent", alpha = 0.8) +
  scale_fill_manual(values = canape_cols) +
  map_theme() +
  theme(legend.title = element_blank())
```

@@

```{r bioregions-make}
# Taxonomic bioregions
a <- ggplot() +
  geom_sf(data = japan, fill = "grey90", color = "black", lwd = 0.25) +
  geom_sf(data = biodiv_ferns_spatial, aes(fill = taxonomic_cluster), color = "transparent", alpha = 0.8) +
  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**(a)** Taxonomic</span>", hjust = "inward", vjust = "inward", color = "white") + 
  labs(fill = "Cluster") +
  scale_fill_manual(values = bioregion_cols) +
  map_theme()

# Phylogenetic bioregions
b <- ggplot() +
  geom_sf(data = japan, fill = "grey90", color = "black", lwd = 0.25) +
  geom_sf(data = biodiv_ferns_spatial, aes(fill = phylo_cluster), color = "transparent", alpha = 0.8) +
  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**(b)** Phylogenetic</span>", hjust = "inward", vjust = "inward", color = "white") + 
  labs(fill = "Cluster") +
  scale_fill_manual(values = bioregion_cols) +
  map_theme() +
  theme(
    axis.text.y = element_blank()
  )

@@

cluster_cols <- carto_pal(8, "Bold")[1:4] %>%
  set_names(1:4)

a <- ggplot(data, aes(x = taxonomic_cluster, y = pd_obs_z)) +
  geom_jitter(aes(color = taxonomic_cluster)) +
  scale_color_manual(values = cluster_cols, breaks = 1:4) +
  geom_boxplot(color = "grey80", fill = "transparent", outlier.shape = NA, alpha = 0.25) +
  labs(
    y = "SES PD",
    x = "Cluster",
    color = "Cluster"
  ) +
  theme_bw() +
  theme(
    panel.grid.minor.x = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.y = element_blank(),
    axis.text.x = element_blank(),
    axis.title.x = element_blank(),
    axis.ticks.x = element_blank()
  )

b <- ggplot(data, aes(x = taxonomic_cluster, y = rpd_obs_z)) +
  geom_jitter(aes(color = taxonomic_cluster)) +
  scale_color_manual(values = cluster_cols, breaks = 1:4) +
  geom_boxplot(color = "grey80", fill = "transparent", outlier.shape = NA, alpha = 0.25) +
  labs(
    y = "SES RPD",
    x = "Cluster",
    color = "Cluster"
  ) +
  theme_bw() +
  theme(
    panel.grid.minor.x = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.y = element_blank(),
    axis.text.x = element_blank(),
    axis.title.x = element_blank(),
    axis.ticks.x = element_blank()
  )

c <- ggplot(data, aes(x = taxonomic_cluster, y = fd_obs_z)) +
  geom_jitter(aes(color = taxonomic_cluster)) +
  scale_color_manual(values = cluster_cols, breaks = 1:4) +
  geom_boxplot(color = "grey80", fill = "transparent", outlier.shape = NA, alpha = 0.25) +
  labs(
    y = "SES FD",
    x = "Cluster",
    color = "Cluster"
  ) +
  theme_bw() +
  theme(
    panel.grid.minor.x = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.y = element_blank(),
    axis.text.x = element_blank(),
    axis.title.x = element_blank(),
    axis.ticks.x = element_blank()
  )

d <- ggplot(data, aes(x = taxonomic_cluster, y = rfd_obs_z)) +
  geom_jitter(aes(color = taxonomic_cluster)) +
  scale_color_manual(values = cluster_cols, breaks = 1:4) +
  geom_boxplot(color = "grey80", fill = "transparent", outlier.shape = NA, alpha = 0.25) +
  labs(
    y = "SES RFD",
    x = "Cluster",
    color = "Cluster"
  ) +
  theme_bw() +
  theme(
    panel.grid.minor.x = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.y = element_blank(),
    axis.title.x = element_blank()
  )

e <- ggplot(data, aes(x = taxonomic_cluster, y = pe_obs_p_upper)) +
  geom_jitter(aes(color = taxonomic_cluster)) +
  scale_color_manual(values = cluster_cols, breaks = 1:4) +
  geom_boxplot(color = "grey80", fill = "transparent", outlier.shape = NA, alpha = 0.25) +
  labs(
    y = "PE *p*-score",
    x = "Cluster",
    color = "Cluster"
  ) +
  theme_bw() +
  theme(
    axis.title.y = element_markdown(),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.y = element_blank(),
  )

@@

```{r conserv-status-make}
# Plot protected areas
conserv_status_plot <- signif_cells_protected_area %>%
  mutate(
    metric = case_when(
      metric == "pd" ~ "PD",
      metric == "fd" ~ "FD",
      metric == "pe" ~ "PE",
      metric == "richness" ~ "Richness",
    ),
    status = str_to_sentence(status)) %>%
  mutate(metric = factor(metric, levels = c("PE", "FD", "PD", "Richness"))) %>%
  ggplot(aes(x = percent_protected, y = metric, fill = status)) +
  geom_col(position = position_dodge()) +
  geom_vline(xintercept = total_percent_protected_medium, color = "grey40", linetype = 3) +
  geom_vline(xintercept = total_percent_protected_high, color = "grey40", linetype = 2) +
  scale_fill_viridis_d() +
  scale_x_continuous(labels = scales::percent) +
  labs(
    x = "Percent protected",
    y = "Biodiversity metric",
    fill = "Protection status"
  ) +
  theme_bw() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    legend.position = "bottom"
  )
```

>>>>

```{r setup, include = FALSE, cache = FALSE}
# Load libraries used for rendering Rmds
library(flextable)
library(kableExtra)
library(ftExtra)
library(scales)
library(patchwork)
library(RColorBrewer)

@@

# Load functions and variables only used when rendering Rmds
source(here::here("R/ms_functions.R"), local = TRUE)
```

```{r load-data}
# Load objects from targets workflow
tar_load(c(
  "repro_data",
  "occ_point_data_ferns",
  "occ_point_data_ferns_unfiltered",
  "biodiv_ferns_spatial",
  "k_taxonomy",
  "k_phylogeny",
  "signif_cells_protected_area",
  "protected_areas",
  "japan_shp",
  "phy_sig_results",
  "binary_sig_results",
  "traits_summary",
  "fern_traits",
  "japan_pterido_tree",
  "japan_fern_tree",
  "ppgi",
  "occ_point_data_ferns",
  "comm_ferns_full",
  "comm_ferns",
  "japan_points_raw",
  "plastome_calibration_dates",
  "plastome_tree",
  "green_list",
  "model_params",
  "lrt_results",
  "aic_env_repro",
  "model_fits",
  "biodiv_ferns_cent",
  "refs_yaml",
  "refs_other_yaml",
  "template_file",
  "csl_file",
  "broad_alignment_list",
  "japan_rbcL",
  "ms_functions"
))
```

@@

Sequences for the plastid, coding *rbcL* gene are available for ca. 97% of the Japanese fern flora [@Ebihara2019b]. 
However, using this dataset alone for phylogeny suffers two drawbacks: it is not densely sampled enough to resolve many internal nodes, and it lacks many lineages that have fossils available for molecular dating. 
Because one goal of the study is to understand the distribution of paleo- vs. neo-endemic areas (defined in units of time), we require an ultrametric tree. 
To obtain a maximally sampled, robust phylogenetic tree, we combined the *rbcL* data of @Ebihara2019b (`r n_japan_fern_taxa` taxa) with a globally sampled plastid dataset including all fern sequences on GenBank for four widely-sequenced plastid genes (*atpA*, *atpB*, *rbcL*, and *rps4*; `r n_ftol_fern_taxa` taxa) and `r n_plastome_genes` other coding, single-copy plastid genes extracted from `r n_plastome_fern_taxa` complete fern plastomes (Nitta et al., in prep.) (`r ape::Ntip(plastome_tree) %>% scales::number(big.mark = ",")` OTUs total, including outgroup taxa). 
We conducted phylogenetic analysis with `r software("IQTREE")` [@Nguyen2015] and molecular dating with `r software("treePL")` [@Smith2012] using `r nrow(plastome_calibration_dates)` fossil calibration points after @Testo2016a, with the exception of treating fossil *Kuylisporites* as belonging to crown *Cyathea*, not *Cyathea* + *Alsophila* [@Loiseau2020]. 
We then trimmed the tree to Japanese taxa only and used this for all further analyses.

@@

```{r plot-prep}
# (color palettes are set in ms_functions.R)

# Define nicely formatted terms for printing variable names
resp_var_lookup <- tribble(
  ~resp_var, ~resp_var_pretty,
  "fd_obs_z", "SES of FD",
  "pd_obs_z", "SES of PD",
  "pe_obs_p_upper", "PE *p*-value",
  "rfd_obs_z", "SES of RFD",
  "richness", "Richness",
  "rpd_obs_z", "SES of RPD",
)

@@

# Crop high-res map of Japan to spatial div results area
japan <- sf::st_crop(japan_shp, sf::st_bbox(biodiv_ferns_spatial))
japan_crs <- sf::st_crs(japan)

# Format biodiv_ferns_spatial data
biodiv_ferns_spatial <-
  biodiv_ferns_spatial %>% 
  # Set CRS
  sf::st_set_crs(japan_crs) %>%
  # Set order of factor levels for plotting
  mutate(
    endem_type = factor(endem_type, levels = c("neo", "paleo", "not significant", "mixed", "super")),
    # can ignore warning about unknown levels, since they don't exist in all variables but that's OK
    across(
      contains("signif"),
      ~fct_relevel(.x, "< 0.01",  "< 0.025", "not significant", "> 0.975", "> 0.99")
    )
  )
```

@@

```{r raw-diversity-make}
# Use thinner coast lines for this plot
# it's smaller, so the grey lines appear relatively thicker
coast_lwd_old <- coast_lwd
coast_lwd <- 0.05

# Richness
a <- ggplot() +
  geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) +
  geom_sf(data = biodiv_ferns_spatial, aes(fill = richness), color = "transparent") +
  geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) +
  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**(a)** Richness</span>", hjust = "inward", vjust = "inward", color = "white") + 
  labs(fill = "No. taxa") +
  scale_fill_scico(palette = "lajolla") + 
  scale_x_continuous(breaks = c(125, 135, 145)) +
  scale_y_continuous(breaks = c(25, 35, 45)) +
  map_theme()

# Raw PD
b <- ggplot() +
  geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) +
  geom_sf(data = biodiv_ferns_spatial, aes(fill = pd_obs), color = "transparent") +
  geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) +
  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**(b)** PD</span>", hjust = "inward", vjust = "inward", color = "white") +
  labs(fill = "Relative\nbranch\nlength") +
  scale_fill_scico(palette = "lajolla") + 
  scale_x_continuous(breaks = c(125, 135, 145)) +
  scale_y_continuous(breaks = c(25, 35, 45)) +
  map_theme() +
  theme(axis.text.y = element_blank())

# Raw FD
c <- ggplot() + 
  geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) +
  geom_sf(data = biodiv_ferns_spatial, aes(fill = fd_obs), color = "transparent") +
  geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) +
  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**(c)** FD</span>", hjust = "inward", vjust = "inward", color = "white") + 
  labs(fill = "Relative\nbranch\nlength") +
  scale_fill_scico(palette = "lajolla") + 
  scale_x_continuous(breaks = c(125, 135, 145)) +
  scale_y_continuous(breaks = c(25, 35, 45)) +
  map_theme() +
  theme(axis.text.y = element_blank())

raw_diversity_plot <- a + b + c +
  plot_layout(nrow = 1)

# Reset coastline width
coast_lwd <- coast_lwd_old
```

@@

```{r rand-diversity-make}
# PD
a <- ggplot() +
  # white base map, no outline
  geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) +
  # colored cells
  geom_sf(data = biodiv_ferns_spatial, aes(fill = pd_signif), color = "transparent") +
  # light grey coastline
  geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) +
  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**(a)** PD</span>", hjust = "inward", vjust = "inward", color = "white") + 
  scale_fill_manual(values = signif_cols) +
  map_theme() +
  theme(
    axis.text.x = element_blank(),
    legend.title = element_blank()
  )

# RPD
b <- ggplot() +
  geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) +
  geom_sf(data = biodiv_ferns_spatial, aes(fill = rpd_signif), color = "transparent") +
  geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) +
  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**(b)** RPD</span>", hjust = "inward", vjust = "inward", color = "white") + 
  labs(fill = "") +
  scale_fill_manual(values = signif_cols) +
  map_theme() +
  theme(
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    legend.title = element_blank()
  )

# FD
c <- ggplot() +
  geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) +
  geom_sf(data = biodiv_ferns_spatial, aes(fill = fd_signif), color = "transparent") +
  geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) +
  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**(c)** FD</span>", hjust = "inward", vjust = "inward", color = "white") + 
  labs(fill = "") +
  scale_fill_manual(values = signif_cols) +
  scale_x_continuous(guide = guide_axis(check.overlap = TRUE)) +
  map_theme() +
  theme(legend.title = element_blank())

# RFD
d <- ggplot() +
  geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) +
  geom_sf(data = biodiv_ferns_spatial, aes(fill = rfd_signif), color = "transparent") +
  geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) +
  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**(d)** RFD</span>", hjust = "inward", vjust = "inward", color = "white") + 
  labs(fill = "") +
  scale_fill_manual(values = signif_cols) +
  scale_x_continuous(guide = guide_axis(check.overlap = TRUE)) +
  map_theme() +
  theme(
    axis.text.y = element_blank(),
    legend.title = element_blank()
  )

@@

# Generate final plot
modelfit_plot_list <- 
  model_fits %>%
  mutate(
    signif_var = str_replace_all(resp_var, "_obs_z", "_signif"),
    resp_var_print = case_when(
      resp_var == "richness" ~ "Richness (no. spp.)",
      TRUE ~ str_to_upper(resp_var) %>% word(1, 1, "_") %>% paste("SES of", .)
    ),
    model_data = list(biodiv_ferns_used_for_model),
    signif_cols = list(signif_cols)) %>%
  left_join(model_subplot_order, by = c("resp_var", "indep_var")) %>%
  arrange(order) %>%
  pmap(plot_fits)

@@

```{r endemism-make}
endemism_plot <- ggplot() + 
  geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) +
  geom_sf(data = biodiv_ferns_spatial, aes(fill = endem_type), color = "transparent") +
  geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) +
  scale_fill_manual(values = canape_cols) +
  map_theme() +
  theme(legend.title = element_blank())
```

@@

```{r bioregions-make}
# Taxonomic bioregions
a <- ggplot() +
  geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) +
  geom_sf(data = biodiv_ferns_spatial, aes(fill = taxonomic_cluster), color = "transparent") +
  geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) +
  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**(a)** Taxonomic</span>", hjust = "inward", vjust = "inward", color = "white") + 
  labs(fill = "Cluster") +
  scale_fill_manual(values = bioregion_cols) +
  map_theme()

# Phylogenetic bioregions
b <- ggplot() +
  geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) +
  geom_sf(data = biodiv_ferns_spatial, aes(fill = phylo_cluster), color = "transparent") +
  geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) +
  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**(b)** Phylogenetic</span>", hjust = "inward", vjust = "inward", color = "white") + 
  labs(fill = "Cluster") +
  scale_fill_manual(values = bioregion_cols) +
  map_theme() +
  theme(axis.text.y = element_blank())

@@

a <- ggplot(data, aes(x = taxonomic_cluster, y = pd_obs_z)) +
  geom_jitter(aes(color = taxonomic_cluster)) +
  scale_color_manual(values = bioregion_cols, breaks = 1:4) +
  geom_boxplot(color = "grey80", fill = "transparent", outlier.shape = NA, alpha = 0.25) +
  labs(
    y = "SES PD",
    x = "Cluster",
    color = "Cluster"
  ) +
  theme_bw() +
  theme(
    panel.grid.minor.x = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.y = element_blank(),
    axis.text.x = element_blank(),
    axis.title.x = element_blank(),
    axis.ticks.x = element_blank()
  )

b <- ggplot(data, aes(x = taxonomic_cluster, y = rpd_obs_z)) +
  geom_jitter(aes(color = taxonomic_cluster)) +
  scale_color_manual(values = bioregion_cols, breaks = 1:4) +
  geom_boxplot(color = "grey80", fill = "transparent", outlier.shape = NA, alpha = 0.25) +
  labs(
    y = "SES RPD",
    x = "Cluster",
    color = "Cluster"
  ) +
  theme_bw() +
  theme(
    panel.grid.minor.x = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.y = element_blank(),
    axis.text.x = element_blank(),
    axis.title.x = element_blank(),
    axis.ticks.x = element_blank()
  )

c <- ggplot(data, aes(x = taxonomic_cluster, y = fd_obs_z)) +
  geom_jitter(aes(color = taxonomic_cluster)) +
  scale_color_manual(values = bioregion_cols, breaks = 1:4) +
  geom_boxplot(color = "grey80", fill = "transparent", outlier.shape = NA, alpha = 0.25) +
  labs(
    y = "SES FD",
    x = "Cluster",
    color = "Cluster"
  ) +
  theme_bw() +
  theme(
    panel.grid.minor.x = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.y = element_blank(),
    axis.text.x = element_blank(),
    axis.title.x = element_blank(),
    axis.ticks.x = element_blank()
  )

d <- ggplot(data, aes(x = taxonomic_cluster, y = rfd_obs_z)) +
  geom_jitter(aes(color = taxonomic_cluster)) +
  scale_color_manual(values = bioregion_cols, breaks = 1:4) +
  geom_boxplot(color = "grey80", fill = "transparent", outlier.shape = NA, alpha = 0.25) +
  labs(
    y = "SES RFD",
    x = "Cluster",
    color = "Cluster"
  ) +
  theme_bw() +
  theme(
    panel.grid.minor.x = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.y = element_blank(),
    axis.title.x = element_blank()
  )

e <- ggplot(data, aes(x = taxonomic_cluster, y = pe_obs_p_upper)) +
  geom_jitter(aes(color = taxonomic_cluster)) +
  scale_color_manual(values = bioregion_cols, breaks = 1:4) +
  geom_boxplot(color = "grey80", fill = "transparent", outlier.shape = NA, alpha = 0.25) +
  labs(
    y = "PE *p*-score",
    x = "Cluster",
    color = "Cluster"
  ) +
  theme_bw() +
  theme(
    axis.title.y = element_markdown(),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.y = element_blank(),
  )

@@

```{r conserv-status-make}
# Plot protected areas
conserv_status_plot <- signif_cells_protected_area %>%
  mutate(
    metric = case_when(
      metric == "pd" ~ "PD",
      metric == "fd" ~ "FD",
      metric == "pe" ~ "PE",
      metric == "richness" ~ "Richness",
    ),
    status = str_to_sentence(status)) %>%
  mutate(metric = factor(metric, levels = c("PE", "FD", "PD", "Richness"))) %>%
  ggplot(aes(x = percent_protected, y = metric, fill = status)) +
  geom_col(position = position_dodge()) +
  geom_vline(xintercept = total_percent_protected_medium, color = "grey40", linetype = 3) +
  geom_vline(xintercept = total_percent_protected_high, color = "grey40", linetype = 2) +
  scale_fill_manual(values = protection_cols) +
  scale_x_continuous(labels = function(x) scales::percent(x, accuracy = 1)) +
  labs(
    x = "Percent protected",
    y = "Biodiversity metric",
    fill = "Protection status"
  ) +
  theme_bw() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    legend.position = "bottom"
  )
```
